Revealing transient strain in geodetic data with Gaussian process regression

### 2017-10-10 RevisedSubmission/hines_hetland_2017_gji.tex
did more

\section{Outlier detection algorithm}
Our outlier detection algorithm is loosely based on the data editing algorithm from \citet{Acheson1975}. Let $\mitbf{d}^*$ denote all $n$ GNSS displacement observations for a single directional component, which have been made at positions and times $\mitbf{P}$.  We model $\mitbf{d}^*$ as a realization of the random vector 
\begin{equation}\label{eq:OutlierDataVec}
\mitbf{d} = \mitbf{G}\mitbf{m} + v(\mitbf{P}) + \mitbf{w},
\end{equation}
where $\mitbf{G}$ and $\mitbf{m}$ are the same as in eq. (\ref{eq:DataVec}), $v$ is a Gaussian process distributed as $\mathcal{GP}(0,C_v)$, and $\mitbf{w}$ is a vector of uncorrelated noise with known standard deviations $\mitbf{\sigma} = [\sigma_1, \sigma_2, ..., \sigma_n]$. The Gaussian process $v$ is intended to describe transient features in the data that cannot be explained by the linear trend or seasonal terms in $\mitbf{G}$. We let the temporal covariance of $v$ be a squared exponential, and we let $v$ be spatially uncorrelated so that,
\begin{equation}
C_v(p,p') = \phi^2\exp\left(\frac{-|t - t'|^2}{2\tau^2}\right) C_w(\vec{x},\vec{x}'),
\end{equation}
where 
\begin{equation}\label{eq:WhiteCov}
C_w(\vec{x},\vec{x}') = 
\begin{cases}
1, &\vec{x} = \vec{x}' \\
0, &\vec{x} \neq \vec{x}'
\end{cases}.
\end{equation}
The spatial covariance of $v$ has little effect on the detected outliers, and so we have assumed that $v$ is spatially uncorrelated for simplicity. Based on our experience, $v$ can reasonably describe most transient features in the data when we set $\phi = 1$ mm and $\tau = 10$ days.

>>>>

\section{Outlier detection algorithm}
Our outlier detection algorithm is loosely based on the data editing algorithm from \citet{Acheson1975}. Let $\mitbf{d}^*$ denote all $n$ GNSS displacement observations for a single directional component, which have been made at positions and times $\mitbf{P}$.  We model $\mitbf{d}^*$ as a realization of the random vector 
\begin{equation}\label{eq:OutlierDataVec}
\mitbf{d} = \mitbf{G}\mitbf{m} + v(\mitbf{P}) + \mitbf{w},
\end{equation}
where $\mitbf{G}$ and $\mitbf{m}$ are the same as in eq. (\ref{eq:DataVec}), $v$ is a Gaussian process distributed as $\mathcal{GP}(0,C_v)$, and $\mitbf{w}$ is a vector of uncorrelated Gaussian noise with known standard deviations $\mitbf{\sigma} = [\sigma_1, \sigma_2, ..., \sigma_n]$. The Gaussian process $v$ is intended to describe transient features in the data that cannot be explained by the linear trend or seasonal terms in $\mitbf{G}$. We let the temporal covariance of $v$ be a squared exponential, and we let $v$ be spatially uncorrelated so that,
\begin{equation}
C_v(p,p') = \phi^2\exp\left(\frac{-|t - t'|^2}{2\tau^2}\right) C_w(\vec{x},\vec{x}'),
\end{equation}
where 
\begin{equation}\label{eq:WhiteCov}
C_w(\vec{x},\vec{x}') = 
\begin{cases}
1, &\vec{x} = \vec{x}' \\
0, &\vec{x} \neq \vec{x}'
\end{cases}.
\end{equation}
The spatial covariance of $v$ has little effect on the detected outliers, and so we have assumed that $v$ is spatially uncorrelated for simplicity. Based on our experience, $v$ can reasonably describe most transient features in the data when we set $\phi = 1$ mm and $\tau = 10$ days.
